---
title: "QAQC All Protocols"
author: "Sophie Dua, Jacob Davison"
date: "2026-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading packages

Setting up R and installing packages (as needed). Make sure to always load appropriate packages.

```{r}
#find working directory
getwd()

#Install packages if you need
# install.packages("tidyverse")
# install.packages("dplyr")
# install.packages("writexl")
# install.packages("EnvStats")
# install.packages("janitor")

#load packages !Very important!
# tidyverse and dplyr have lots of useful functions for data cleaning
library(tidyverse)
library(dplyr)
# writexl is used to create excel files from within R
library(writexl)
# knitr is used to create output files from R Markdown
library(knitr)
# EnvStats is needed for the rosnerTest() function
library(EnvStats)
# janitor needed for get_dupes()
library(janitor)
```

## Loading data: Paste all file paths here!

Load all the appropriate .csv files that we want to QAQC. Title each protocol dataset and adjust file paths.

```{r}
#load data 
Fuels_FWD<- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Surface Fuels - Fine_XPT.csv")

Fuels_1000HR<- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Surface Fuels - 1000Hr_XPT.csv")

Fuels_LittDuff<- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Surface Fuels - Duff_Litter_XPT.csv")

Post_Burn <- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Post Burn Severity (metric)_XPT.csv")

Seedlings <- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Density - Quadrats (metric)_XPT.csv")

Trees<- read.csv("/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/data/Trees - Individuals (metric)_XPT.csv")

```

## Clean up the data

This code cleans up the data a little more by making the date/year and plot info more readable.

```{r}
# Clean up data frame
df_names <- c("Fuels_FWD", "Fuels_1000HR", "Fuels_LittDuff", "Post_Burn", "Seedlings", "Trees")

for (name in df_names)
  
{current_df <- get(name) %>%
  # Separate Date column into Date and Time columns
  separate(Date, sep = " ", into = c("Date", "Time")) %>%
  separate(Date, sep = "/", into = c("Month", "Day", "Year"), remove = FALSE) %>% 
  select(!c("Month", "Day")) %>% 
  # Separate MacroPlotName to get Monitoring type and plot number
  separate(
    MacroPlot.Name,
    into = c("Drop", "MonitoringType", "Plot"),
    sep = ":",
    remove = FALSE
  ) %>%
  #delete any blank spaces in our character variables using trimws()
  mutate(across(where(is.character), ~na_if(trimws(.x), ""))) %>%
  
  select(-Drop)

# Ensure blanks in Visited column are filled in as NA
current_df$Visited[current_df$Visited==""] <- NA

# Convert to Date with explicit format
# Change %m/%d/%Y to whatever your CSV actually uses. For FFI it outputs as %m/%d/%Y
current_df$Date <- as.Date(current_df$Date, format = "%m/%d/%Y")

assign(name, current_df)
}
```

Separate header data from other data for each protocol.

```{r}
# Filter for only data into a new data frames
FuelsFWD_data <- Fuels_FWD %>% filter(is.na(Visited))
Fuels1000_data <- Fuels_1000HR %>% filter(is.na(Visited))
Fuels_LittDuff_Data <- Fuels_LittDuff %>% filter(is.na(Visited))
Post_Burn_data <- Post_Burn %>% filter(is.na(Visited))
Seedlings_Data <- Seedlings %>% filter(is.na(Visited))
Trees_data <- Trees %>%filter(is.na(Visited))


# Filter for only headers in a different data frames
FuelsFWD_header <- Fuels_FWD %>% filter(!is.na(Visited))
Fuels1000_header <- Fuels_1000HR %>% filter(!is.na(Visited))
Fuels_LittDuff_Header <- Fuels_LittDuff %>% filter(!is.na(Visited))
Post_Burn_header <- Post_Burn %>% filter(!is.na(Visited))
Seedlings_Header <- Seedlings %>% filter(!is.na(Visited))
Trees_header <-Trees %>%filter(!is.na(Visited))

```

## Create a blank error log

This error log data frame will be filled in as we run the qa/qc function unless no errors are present.

```{r}
# Blank data frame for no errors
errors_blank <- data.frame("SavedQuery" = "",
                           "MacroPlot.Name" = "",
                           "Date" = as.Date(NA),
                           "Error" = "No Error",
                           "Comment" = "")
```

## Create QA/QC function

This function needs four variable inputs to operate: data, query, query_message, and values_check. The output of this function is a complete error log.

Values_check should be a string of T/F statements. A FALSE value will be read as an error and flagged. An NA value will be set to FALSE and also flagged as an error. Errors are also consolidated by plot and date to make the error logs more readable.

```{r}

# Create error group. Duplicate errors will be grouped by these columns.
errors_group <- c("SavedQuery", "MacroPlot.Name", "Date", "Error")

# QAQC function
qaqc <- function(data, query, query_message, values_check) {
  
  # Check if there is data to QAQC
  if(nrow(data) == 0) {
    # If there is no data, default to "No Error" data frame
    errors <- errors_blank %>%
      mutate("SavedQuery" = query,
             "Error" = "No Error")
    # If there is data, perform error check
  } else {
    
    # Use relevant dataset to look for errors
    errors <- data %>%
      # Add columns relevant to error checking
      # Populate the "SavedQuery" column with the query name
      mutate("SavedQuery" = query,
             # Populate the "Error" column with the query message and relevant data
             "Error" = paste(query_message, "=", values_data),
            ) %>%
      
      # Filter for data which is "false" (data does not match valid conditions)
      # NA values are treated as "false" (missing data is not valid)
      filter(!(values_check %>% replace_na(FALSE))) %>%
      # Select relevant columns to view errors
      select(colnames(errors_blank))
    
    # Next, check if there are duplicate errors
    errors_temp <- errors %>%
      get_dupes(all_of(errors_group))
    # Merge duplicate errors
    errors <- unique(merge(errors, errors_temp, all = T))
    # If duplicate errors exist, add number of duplicates to error message
    errors <- errors %>%
      mutate(Error = ifelse(is.na(dupe_count), Error, paste0("(x", dupe_count, ") ", Error))) %>%
      select(!"dupe_count")
  }
  
  # Check if there are no errors
  if (nrow(errors) == 0) {
    # If there are no errors, default to "No Error" data frame
    errors <- errors_blank %>%
      mutate(SavedQuery = query,
             Error = "No Error")
    # If there are errors, keep error log you just created
  } else {
    errors <- errors
  }}
```

## Begin checking for errors protocol by protocol...

## FUELS FWD ERROR CHECKS

#### Header Metadata:

Number of Transects (NumTran) = 4 and Transect Number (TransNum) = 1, 2, 3, or 4

```{r}

# NumTran
# Set parameters  for the QAQC function

# Load the data frame containing the "FuelsFWD" header information
data <- FuelsFWD_header
# Define a descriptive string for the query related to the number of transects
query <- "Fuel FWD Header NumTran"
# Set a message to indicate what aspect is being checked
query_message <- "Number of Transects" 
# Extract the "NumTran" column from the data frame into a variable for checking
values_data <- data$NumTran
# Define the valid number of transects expected in the data
values_valid <- 4
# create a logical vector (T,T,F,T...etc.)indicating which rows meet the valid transect condition
values_check <- values_data == values_valid

# Identify errors by calling the QAQC function and store the results
errors_FuelsFWD_NumTran <- qaqc(data, query, query_message, values_check)


# TranNum
#Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

#Identify errors
errors_FuelsFWD_TransNum <- qaqc(data, query, query_message, values_check)

```

Transect length (TranLen): 1Hr = 6, 10Hr = 6, 100Hr = 12

```{r}
# Start with 1HR transect length 
# Set parameters
data <- FuelsFWD_header
query <- "Fuel FWD Header 1HR TranLen" 
query_message <- "1HR Transect Length " 
values_data <- data$OneHrTranLen 
values_valid <- 6 
values_check <- values_data == values_valid

#Identify errors
errors_FuelsFWD_1HR_TranLen <- qaqc(data, query, query_message, values_check)


# 10HR transect length 
# Set parameters
data <- FuelsFWD_header
query <- "Fuel FWD Header 10HR TranLen" 
query_message <- "10HR Transect Length " 
values_data <- data$TenHrTranLen 
values_valid <- 6 
values_check <- values_data == values_valid

#Identify errors
errors_FuelsFWD_10HR_TranLen <- qaqc(data, query, query_message, values_check)


# 100HR transect length
# Set parametrs
data <- FuelsFWD_header
query <- "Fuel FWD Header 100HR TranLen" 
query_message <- "100HR Transect Length " 
values_data <- data$HunHrTranLen 
values_valid <- 12 
values_check <- values_data == values_valid

#Identify errors
errors_FuelsFWD_100HR_TranLen <- qaqc(data, query, query_message, values_check)

```

Slope values are reasonable.

```{r}
# Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- (values_data %in% values_valid) %>% replace_na(TRUE)

# Identify errors
errors_FuelsFWD_Slope <- qaqc(data, query, query_message, values_check)

```

Azimuth values are between 0 and 360.

```{r}
# Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values Azimuth"
query_message <- "Azimuth"
values_data <- data$Azimuth
values_valid <- seq(0, 360)
values_check <- (values_data %in% values_valid) %>% replace_na(TRUE)

#Identify errors
errors_FuelsFWD_Azimuth <- qaqc(data, query, query_message, values_check)

```

Check that fuels counts are postitive integer values, and reasonable using a Rosner's outliers test.

1HR:

```{r}

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_FWD %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(OneHr))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$OneHr)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_OneHr <- min(outliers) - 1

# Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values OneHr"
query_message <- "OneHr"
values_data <- data$OneHr
values_valid <- seq(0, max_OneHr, by = 1)
values_check <- !is.na(values_data) & values_data >= 0 & values_data <= max_OneHr


# Identify errors
errors_FuelsFWD_OneHr <- qaqc(data, query, query_message, values_check)
```

10HR:

```{r}

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_FWD %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(TenHr))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$TenHr)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_TenHr <- min(outliers) - 1

# Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values TenHr"
query_message <- "TenHr"
values_data <- data$TenHr
values_valid <- seq(0, max_TenHr, by = 1)
values_check <- !is.na(values_data) & values_data >= 0 & values_data <= max_TenHr


# Identify errors
errors_FuelsFWD_TenHr <- qaqc(data, query, query_message, values_check)
```

100HR:

```{r}

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_FWD %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(HunHr))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$HunHr)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_HunHr <- min(outliers) - 1

# Set parameters
data <- FuelsFWD_data
query <- "Fuel FWD Values HunHr"
query_message <- "HunHr"
values_data <- data$HunHr
values_valid <- seq(0, max_HunHr, by = 1)
values_check <- !is.na(values_data) & values_data >= 0 & values_data <= max_HunHr

# Identify errors
errors_FuelsFWD_HunHr <- qaqc(data, query, query_message, values_check)
```

## Fuels FWD Error Log

```{r}
# Save to master error list
errors_FuelsFWD <- rbind(errors_FuelsFWD_TransNum, errors_FuelsFWD_NumTran, errors_FuelsFWD_1HR_TranLen, errors_FuelsFWD_10HR_TranLen, errors_FuelsFWD_100HR_TranLen, errors_FuelsFWD_Slope, errors_FuelsFWD_Azimuth, errors_FuelsFWD_OneHr, errors_FuelsFWD_TenHr, errors_FuelsFWD_HunHr)
```

```{r}
#remove all the error data frames
remove(errors_FuelsFWD_TransNum, errors_FuelsFWD_NumTran, errors_FuelsFWD_1HR_TranLen, errors_FuelsFWD_10HR_TranLen, errors_FuelsFWD_100HR_TranLen, errors_FuelsFWD_Slope, errors_FuelsFWD_Azimuth, errors_FuelsFWD_OneHr, errors_FuelsFWD_TenHr, errors_FuelsFWD_HunHr)
```

## FUELS 1000S ERROR CHECKS

#### Header Metadata

Number of transects (NumTran) = 4

```{r}
# Set parameters  for the QAQC function

# Load the data frame containing the "1000 Fuel" header information
data <- Fuels1000_header
# Define a descriptive string for the query related to the number of transects
query <- "Fuel 1000 Header NumTran"
# Set a message to indicate what aspect is being checked
query_message <- "Number of Transects" 
# Extract the "NumTran" column from the data frame into a variable for checking
values_data <- data$NumTran
# Define the valid number of transects expected in the data
values_valid <- 4
# create a logical vector (T,T,F,T...etc.)indicating which rows meet the valid transect condition
values_check <- values_data == values_valid

# Identify errors by calling the QAQC function and store the results
errors_Fuels1000_NumTran <- qaqc(data, query, query_message, values_check)

```

Transect length (TranLen) = 50

```{r}
# Set parameters 
data <- Fuels1000_header
query <- "Fuel 1000 Header TranLen" 
query_message <- "Transect Length" 
values_data <- data$TranLen 
values_valid <- 50  
values_check <- values_data == values_valid


# Identify errors
errors_Fuels1000_TranLen <- qaqc(data, query, query_message, values_check)

```

#### 1000s Fuels Data

Decay class (DecayCl) = 3 or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 DecayClass"
query_message <- "Decay Class"
values_data <- data$DecayCl
values_valid <- c(3,4)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_DecayCl <- qaqc(data, query, query_message, values_check)

```

Transect number (TransNum) = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_Fuels1000_TransNum <- qaqc(data, query, query_message, values_check)

```

Slope values are reasonable.

```{r}
# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Slope"
query_message <- "Slope"
values_data <- data$Slope
values_valid <- seq(0, 90, by = 1)
values_check <- (values_data %in% values_valid) %>% replace_na(TRUE)

# Identify errors
errors_Fuels1000_Slope <- qaqc(data, query, query_message, values_check)
```

Diameter values are reasonable: not an outlier based on a Rosner's test and ≥ 3.0.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_1000HR %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(Dia))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$Dia)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Dia <- min(outliers) - 1

# Set parameters
data <- Fuels1000_data
query <- "Fuel 1000 Values Dia"
query_message <- "Diameter"
values_data <- data$Dia
values_valid <- seq(3, max_Dia)
values_check <- !is.na(values_data) & values_data >= 3 & values_data <= max_Dia

# Identify errors
errors_Fuels1000_Dia <- qaqc(data, query, query_message, values_check)
```

## Fuels 1000s Error Log

```{r}
# Save to master error list
errors_Fuels1000 <- rbind(errors_Fuels1000_DecayCl, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)
```

```{r}
#remove all the error data frames
remove(errors_Fuels1000_DecayCl, errors_Fuels1000_NumTran, errors_Fuels1000_TranLen, errors_Fuels1000_TransNum, errors_Fuels1000_Slope, errors_Fuels1000_Dia)
```

## LITTER & DUFF ERROR CHECKS

#### Header Metadata:

Number of Transects = 4

```{r}
# Load the data frame containing the "Fuels_LittDuff" header information
data <- Fuels_LittDuff_Header
# Define a descriptive string for the query related to the number of transects
query <- "Fuel LittDuff Header NumTran"
# Set a message to indicate what aspect is being checked
query_message <- "Number of Transects" 
# Extract the "NumTran" column from the data frame into a variable for checking
values_data <- data$NumTran
# Define the valid number of transects expected in the data
values_valid <- 4
# create a logical vector indicating which rows meet the valid transect condition
values_check <- trimws(values_data) == values_valid

# Identify errors by calling the QAQC function and store the results
errors_Fuels_LittDuff_NumTran <- qaqc(data, query, query_message, values_check)
```

Transect Number = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Fuels_LittDuff_Data
query <- "Fuel LittDuff Values TransNum"
query_message <- "Transect Number"
values_data <- data$Transect
values_valid <- c(1, 2, 3, 4)
values_check <- trimws(values_data) %in% values_valid

# Identify errors
errors_Fuels_LittDuff_TransNum <- qaqc(data, query, query_message, values_check)
```

Sample Location = 1, 5, 10, 15, 20, 25, 30, 35, 40, 45

```{r}
#Set parameters
data <- Fuels_LittDuff_Data
query <- "Sample Location on Transect"
query_message <- "Sample location"
values_data <- data$SampLoc
values_valid <- c(1, 5, 10, 15, 20, 25, 30, 35, 40, 45)
values_check <- (values_data) %in% values_valid

#Identify errors
errors_Fuels_LittDuff_SampLoc <- qaqc(data, query, query_message, values_check)

```

Also flag plots where zero values are found for both Duff and Litt at Sample locations 5, 15, 25, 35. These zeros are likely errors from pre-1992 when the protocol was changed and should be NA.

```{r}
#Set parameters
data <- Fuels_LittDuff_Data
query <- "Fuel LittDuff SampLoc; False Zeros"
query_message <- paste0("Litter=", data$LittDep, " Duff=", data$DuffDep, " at SampLoc")
values_data <- data$SampLoc
values_check <- data %>%
  group_by(Plot, Date) %>%
  mutate(
    flag =
      all(c(5,15,25,35) %in% SampLoc) &
      all(
        DuffDep[SampLoc %in% c(5,15,25,35)] == 0 &
        LittDep[SampLoc %in% c(5,15,25,35)] == 0
      )
  ) %>%
  ungroup() %>%
  mutate(flag = !flag) %>%
  pull(flag)


# Identify errors
errors_Fuels_LittDuff_SampLoc_zeros <- qaqc(data, query, query_message, values_check)

```

Litter values are \>= 0 and flags outliers using a Rosner's outlier test.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_LittDuff_Data %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(LittDep))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$LittDep)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Litt <- min(outliers) - 1

# Set parameters
data <- Fuels_LittDuff_Data
query <- "Fuel Litt Values"
query_message <- paste0("Litter value at ", data$SampLoc,"ft")
values_data <- data$LittDep
values_valid <- seq(0, max_Litt)
values_check <- !is.na(values_data) & values_data >= 0 & values_data <= max_Litt

# Identify errors
errors_Fuels_Litt_Outlier <- qaqc(data, query, query_message, values_check)
```

Duff values are \>= 0 and flags outliers using a Rosner's outlier test.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Fuels_LittDuff_Data %>%
  filter(is.na(Visited)) %>% 
  filter(!is.na(DuffDep))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$DuffDep)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Duff <- min(outliers) - 1

# Set parameters
data <- Fuels_LittDuff_Data
query <- "Fuel Duff Values"
query_message <- paste0("Duff value at ", data$SampLoc,"ft")
values_data <- data$DuffDep
values_valid <- seq(0, max_Duff)
values_check <- !is.na(values_data) & values_data >= 0 & values_data <= max_Duff

# Identify errors
errors_Fuels_Duff_Outlier <- qaqc(data, query, query_message, values_check)
```

## Litter & Duff Error Log

```{r}
# Save to master error list
errors_Fuels_DuffLitt <- rbind(errors_Fuels_Duff_Outlier, errors_Fuels_Litt_Outlier, errors_Fuels_LittDuff_NumTran, errors_Fuels_LittDuff_SampLoc, errors_Fuels_LittDuff_SampLoc_zeros, errors_Fuels_LittDuff_TransNum)

#remove all the error data frames
remove(errors_Fuels_Duff_Outlier, errors_Fuels_Litt_Outlier, errors_Fuels_LittDuff_NumTran, errors_Fuels_LittDuff_SampLoc, errors_Fuels_LittDuff_SampLoc_zeros, errors_Fuels_LittDuff_TransNum)

```

## POST BURN ERROR CHECKS

#### Header Metadata:

Number of transects (NumTran) = 4

```{r}
# Set parameters  for the QAQC function

# Load the data frame containing the "Post Burn" header information
data <- Post_Burn_header
# Define a descriptive string for the query related to the number of transects
query <- "Post Burn Header NumTran"
# Set a message to indicate what aspect is being checked
query_message <- "Number of Transects" 
# Extract the "NumTran" column from the data frame into a variable for checking
values_data <- data$NumTran
# Define the valid number of transects expected in the data
values_valid <- 4
# create a logical vector indicating which rows meet the valid transect condition
values_check <- trimws(values_data) == values_valid

# Identify errors by calling the QAQC function and store the results
errors_PostBurn_NumTran <- qaqc(data, query, query_message, values_check)

```

Transect length (TranLen) = 50ft

```{r}
# Set parameters 
data <- Post_Burn_header
query <- "Post Burn Header TranLen" 
query_message <- "Transect Length" 
values_data <- as.numeric(data$TranLen)
values_valid <- c(50) 
values_check <- values_data %in% values_valid


# Identify errors
errors_PostBurn_TranLen <- qaqc(data, query, query_message, values_check)

```

Transect number (TransNum) = 1, 2, 3, or 4

```{r}
# Set parameters
data <- Post_Burn_data
query <- "Post Burn Values TransNum"
query_message <- "Transect Number"
values_data <- as.numeric(data$Transect)
values_valid <- c(1, 2, 3, 4)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_TransNum <- qaqc(data, query, query_message, values_check)

```

Number of points on the transect = 10

```{r}
# Set parameters 
data <- Post_Burn_header
query <- "Post Burn Header NumPtsTran" 
query_message <- "Number of Points on Transect" 
values_data <- as.numeric(data$NumPtsTran)
values_valid <- 10 
values_check <- values_data == values_valid


# Identify errors
errors_PostBurn_NumPtsTran <- qaqc(data, query, query_message, values_check)

```

Tape distance = 1, 5, 10, 15, 20, 25, 30, 35, or 40

```{r}
# Set parameters
data <- Post_Burn_data
query <- "Post Burn Values TapeDist"
query_message <- "Tape Distance"
values_data <- as.numeric(data$TapeDist)
values_valid <- c(1, 5, 10, 15, 20, 25, 30, 35, 40, 45)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_TapeDist <- qaqc(data, query, query_message, values_check)

```

#### Post Burn Data

Check that Veg = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- Post_Burn_data
query <- "Post Burn Values Veg"
query_message <- "Veg"
values_data <- as.numeric(data$Veg)
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Veg <- qaqc(data, query, query_message, values_check)

```

Check that Sub = 0, 1, 2, 3, 4, or 5

```{r}
# Set parameters
data <- Post_Burn_data
query <- "Post Burn Values Sub"
query_message <- "Sub"
values_data <- as.numeric(data$Sub)
values_valid <- c(0, 1, 2, 3, 4, 5)
values_check <- values_data %in% values_valid

# Identify errors
errors_PostBurn_Sub <- qaqc(data, query, query_message, values_check)

```

## Post Burn Error Log

```{r}
# Save to master error list
errors_PostBurn <- rbind(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_TransNum, errors_PostBurn_NumPtsTran, errors_PostBurn_TapeDist, errors_PostBurn_Veg, errors_PostBurn_Sub)
```

```{r}
#remove all the error data frames
remove(errors_PostBurn_NumTran, errors_PostBurn_TranLen, errors_PostBurn_TransNum, errors_PostBurn_NumPtsTran, errors_PostBurn_TapeDist, errors_PostBurn_Veg, errors_PostBurn_Sub)
```

## SEEDLINGS ERROR CHECKS

#### Header Metadata:

Number of Transects = 1

```{r}
# Set parameters  for the QAQC function

# Load the data frame containing the "Seedlings" header information
data <- Seedlings_Header
# Define a descriptive string for the query related to the number of transects
query <- "Seedlings Transect"
# Set a message to indicate what aspect is being checked
query_message <- "Transect Number" 
# Extract the "NumTran" column from the data frame into a variable for checking
values_data <- data$NumTran
# Define the valid transect number expected in the data
values_valid <- 1
# create a logical vector (T,T,F,T...etc.)indicating which rows meet the valid transect condition
values_check <- values_data == values_valid

# Identify errors by calling the QAQC function and store the results
errors_Seedlings_Transect <- qaqc(data, query, query_message, values_check)
```

Quadrat Number = 1

```{r}
# Set parameters
data <- Seedlings_Header
query <- "Seedlings Quadrat"
query_message <- "Quadrat Number" 
values_data <- data$NumQuadTran
values_valid <- 1
values_check <- values_data == values_valid

#Identify errors
errors_Seedlings_Quadrat <- qaqc(data, query, query_message, values_check)
```

Quadrat Length = 10 or 25 and is not the same value as Quadrat Width

```{r}
#Set parameters
data <- Seedlings_Header
query <- "Quadrat Length"
query_message <- "Length of Area Sampled"
values_data <- data$QuadLen
values_valid <- c(10, 25)
values_check <-
  values_data %in% values_valid & 
  !(values_data == data$QuadWid)

#Identify errors
errors_Seedlings_length <- qaqc(data, query, query_message, values_check)
```

Quadrat Width = 10 or 25

```{r}
#Set parameters
data <- Seedlings_Header
query <- "Quadrat Width"
query_message <- "Width of Area Sampled"
values_data <- data$QuadWid
values_valid <- c(10, 25)
values_check <- values_data %in% values_valid

#Identify errors
errors_Seedlings_width <- qaqc(data, query, query_message, values_check)
```

Quadrat Area = 250

```{r}
#Set parameters
data <- Seedlings_Header
query <- "Quadrat Area"
query_message <- "Area sampled"
values_data <- data$Area
values_valid <- 250
values_check <- values_data == values_valid

#Identify errors
errors_Seedlings_area <- qaqc(data, query, query_message, values_check)
```

Seedling Count error = 9999

Counts above 10,000 were auto-listed as 9999 in FFI at one time. Essentially, FFI was not able to compute a value higher than that, so values of 9999 should be flagged

```{r}
#Set parameters
data <- Seedlings_Data
query <- "Seedling Count"
query_message <- "Seedling Count 9999 Error"
values_data <- as.numeric(data$Count)
values_valid <-  values_data >= 0 &
                values_data != 9999 &
                values_data == floor(values_data)
values_check <- values_valid

#Identify errors
errors_Seedlings_count <- qaqc(data, query, query_message, values_check)
```

Seedling Height Class = 0.15, 0.3, 0.6, 1, or 2

```{r}
#Set parameters
data <- Seedlings_Data
query <- "Size Height"
query_message <- "Height Class of Seedlings"
values_data <- data$Height
values_valid <- c(0.15, 0.3, 0.6, 1, 2)
values_check <- values_data %in% values_valid

#Identify errors
errors_Seedling_height <- qaqc(data, query, query_message, values_check)
```

## Seedlings Error Log

```{r}
# Save to master error list
errors_Seedlings <- rbind(errors_Seedling_height, errors_Seedlings_area, errors_Seedlings_count, errors_Seedlings_Quadrat, errors_Seedlings_Transect, errors_Seedlings_length, errors_Seedlings_width)

#remove all the error data frames
remove(errors_Seedling_height, errors_Seedlings_area, errors_Seedlings_count, errors_Seedlings_Quadrat, errors_Seedlings_Transect, errors_Seedlings_length, errors_Seedlings_width)
```

## TREES ERROR CHECKS

#### Header Metadata:

Macroplot size = 0.1

```{r}
# Set parameters  for the QAQC function

# Load the data frame containing the "Trees" header information
data <- Trees_header
# Define a descriptive string for the query related to the MacroPlot size
query <- "Trees Header MacroPlot Size" 
# Set a message to indicate what aspect is being checked
query_message <- "MacroPlot size" 
# Extract the "MacroPlotSize" column from the data frame into a variable for checking
values_data <- as.numeric(data$MacroPlotSize)
# Define the valid size expected in the data
values_valid <- 0.1
# create a logical vector (T,T,F,T...etc.)indicating which rows meet the valid size condition
values_check <- values_data == values_valid

# Identify errors by calling the QAQC function and store the results
errors_Trees_MacroPlotSize <- qaqc(data, query, query_message, values_check)

```

Snagplot size = 0.1

```{r}

# Set parameters
data <- Trees_header
query <- "Trees Header SnagPlot Size" 
query_message <- "SnagPlot size" 
values_data <- as.numeric(data$SnagPlotSize)
values_valid <- 0.1
values_check <- values_data == values_valid

#Identify errors
errors_Trees_SnagPlotSize <- qaqc(data, query, query_message, values_check)

```

Break point diameter = 15

```{r}

#Set parameters
data <- Trees_header
query <- "Trees Header Break Point Diameter" 
query_message <- "Break Point Diameter" 
values_data <- as.numeric(data$BrkPntDia)
values_valid <- 15
values_check <- values_data == values_valid

#Identify errors
errors_Trees_BrkPntDia <- qaqc(data, query, query_message, values_check)
```

Field team is filled in

```{r}

#Set parameters
data <- Trees_header
query <- "Trees Header Field Team" 
query_message <- "Field Team" 
values_data <- data$FieldTeam
values_valid <- !is.na(values_data)
values_check <- values_valid

#Identify errors
errors_Trees_FieldTeam <- qaqc(data, query, query_message, values_check)

```

#### Tree data:

```{r}
#First defign dead and live CrwnCl so we can sort through them
dead_classes <- c("DD", "BAD", "BBD", "CUS", "CS", "RS", "LBS")
live_classes <- c("D", "C", "I", "SC", "O")
```

Quarter = 1, 2, 3, 4

```{r}

#Set parameters
data <- Trees_data
query <- "Valid Quarter Numbers"
query_message <- paste0("Tag #", data$TagNo,": Quarter Number")
values_data <- data$QTR
values_valid <- c(1, 2, 3, 4)
values_check <- (values_data %in% values_valid)

#Identify errors
errors_Trees_QTRNum <- qaqc(data, query, query_message, values_check)
```

Tree Tag is a positive integer

```{r}

#Set parameters
data <- Trees_data
query <- "Tree Tag"
query_message <- "Tree Tag Value"
values_data <- data$TagNo
values_valid <- seq(1, 1e6, by = 1)
values_check <- (values_data %in% values_valid) 

#Identify errors
errors_Trees_PosTag <- qaqc(data, query, query_message, values_check)

```

Tree status:

```{r}

#Set parameters
data <- Trees_data
query <- "Tree Status"
query_message <- "Tree Status"
values_data <- data$Status
values_check <- values_data %in% c("L", "D") 

#Identify errors
errors_Trees_Status <- qaqc(data, query, query_message, values_check)

```

DBH values are valid. Here we are using a Rosner's outlier test to flag outliers as well as flagging values that are \<0 and have a dead Crown class: DD, BBD, or CUS. We also are separating out SEGIs and running them in a seperate outliers test.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

#Separate SEGI and nonSEGI trees
XSEGI_data <- Trees_data %>%
  filter(Species.Symbol != "SEGI1")

SEGI_data <- Trees_data %>%
  filter(Species.Symbol == "SEGI1")


#Rosner's for non-SEGIs
# Filter all data to remove NA values for outlier test
data_temp <- XSEGI_data %>%
  filter(!is.na(DBH)) %>%
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$DBH)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_DBH_XSEGI <- min(outliers) - 1

#Rosner's for only SEGIs
# Filter all data to remove NA values for outlier test
data_temp <- SEGI_data %>%
  filter(!is.na(DBH)) %>%
  filter(!CrwnCl %in% c("DD", "BBD", "CUS"))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$DBH)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_DBH_SEGI <- min(outliers) - 1


# Set parameters for non SEGIs
data <- XSEGI_data
query <- "DBH"
query_message <- case_when(
  is.na(data$DBH) ~ "DBH",
  !is.na(data$DBH) & data$DBH > max_DBH_XSEGI ~ paste0("Tag #", data$TagNo, ": DBH (outlier)"),
  data$CrwnCl %in% c("DD", "CUS", "BBD") ~ paste0("Tag #", data$TagNo, ": DBH (", data$CrwnCl, ")"),
  TRUE ~ paste0("Tag #", data$TagNo, ": DBH"))
values_data <- data$DBH
values_check <- (
  # Live trees: DBH valid
  data$CrwnCl %in% live_classes & values_data > 0 & values_data <= max_DBH_XSEGI) |
(
  # Dead trees: DBH valid
  !(data$CrwnCl %in% c("DD", "CUS", "BBD")) & values_data > 0 & values_data <= max_DBH_XSEGI) |
(
  # Dead trees DD, CUS, BBD: NA allowed
  data$CrwnCl %in% c("DD", "CUS", "BBD") & is.na(values_data))


# Identify errors for non SEGIs
errors_Trees_DBH_XSEGI <- qaqc(data, query, query_message, values_check)


# Set parameters for SEGIs
data <- SEGI_data
query <- "DBH"
query_message <- case_when(
  is.na(data$DBH) ~ "DBH",
  !is.na(data$DBH) & data$DBH > max_DBH_SEGI ~ paste0("Tag #", data$TagNo, ": DBH (outlier)"),
  data$CrwnCl %in% c("DD", "CUS", "BBD") ~ paste0("Tag #", data$TagNo, ": DBH (", data$CrwnCl, ")"),
  TRUE ~ paste0("Tag #", data$TagNo, ": DBH"))
values_data <- data$DBH
values_check <- (
  # Live trees: DBH valid
  data$CrwnCl %in% live_classes & values_data > 0 & values_data <= max_DBH_SEGI) |
(
  # Dead trees: DBH valid
  !(data$CrwnCl %in% c("DD", "CUS", "BBD")) & values_data > 0 & values_data <= max_DBH_SEGI) |
(
  # Dead trees DD, CUS, BBD: NA allowed
  data$CrwnCl %in% c("DD", "CUS", "BBD") & is.na(values_data))


# Identify errors for SEGIS
errors_Trees_DBH_SEGI <- qaqc(data, query, query_message, values_check)

# Combine DBH error logs
errors_Trees_DBH <- rbind(errors_Trees_DBH_XSEGI, errors_Trees_DBH_SEGI)

# Remove combined logs
remove(errors_Trees_DBH_XSEGI, errors_Trees_DBH_SEGI)
```

Height values are valid. Again using a Rosner's outlier test and flagging negative numbers.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Trees_data %>%
  filter(!is.na(Ht))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$Ht)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_Ht <- min(outliers) - 1

# Set parameters
data <- Trees_data
query <- "Trees Height "
query_message <- case_when(
  is.na(data$Ht) ~ "Height",
  !is.na(data$Ht) & data$Ht > max_Ht ~ paste0("Tag #", data$TagNo, ": Height (outlier)"),
  data$CrwnCl == "DD" ~ paste0("Tag #", data$TagNo, ": Height (", data$CrwnCl, ")"),
  TRUE ~ paste0("Tag #", data$TagNo, ": Height")
)
values_data <- data$Ht
values_check <- (
  # Valid Heights
  !(data$CrwnCl == "DD") & values_data > 0 & values_data <= max_Ht) |
(
  # Dead and Down trees DD: NA allowed
  data$CrwnCl == "DD" & is.na(values_data))


# Identify errors
errors_Trees_Ht <- qaqc(data, query, query_message, values_check)
```

Live crown base height values are valid. Rosner's test for outliers, flagging negative values, flagging values that are larger than the height of the tree.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Trees_data %>%
  filter(!is.na(LiCrBHt))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$LiCrBHt)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_LiCrBHt <- min(outliers) - 1

# Set parameters
data <- Trees_data
query <- "Live Crown Base Height"
query_message <- case_when(
  is.na(data$LiCrBHt) ~ "Live Crown Base Height",
  !is.na(data$LiCrBHt) & data$LiCrBHt > max_LiCrBHt ~ paste0("Tag #", data$TagNo, ": Live Crown Base Height (outlier)"),
  !is.na(data$LiCrBHt) & data$LiCrBHt > data$Ht ~ paste0("Tag #", data$TagNo, ": Live Crown Base Height > Height"),
  data$CrwnCl %in% dead_classes ~ paste0("Tag #", data$TagNo, ": Live Crown Base Height (", data$CrwnCl, ")"),
  TRUE ~ paste0("Tag #", data$TagNo, ": Live Crown Base Height")
)
values_data <- data$LiCrBHt
values_check <- (
  # Live trees: must have valid value
  data$CrwnCl %in% live_classes &
  values_data >= 0 &
  values_data <= max_LiCrBHt &
  values_data <= data$Ht
) |
(
  # Dead trees: must be NA
  data$CrwnCl %in% dead_classes &
  is.na(values_data)
)


# Identify errors
errors_Trees_LiCrBHt <- qaqc(data, query, query_message, values_check)
```

Crown class values are valid. Status must match CPCs:

Live trees = live classes

Dead trees = dead classes

```{r}

#Set parameters
data <- Trees_data
query <- "Valid Crown Class"
query_message <- case_when(
   (data$Status == "D" & data$CrwnCl %in% live_classes) | (data$Status == "L" & data$CrwnCl %in% dead_classes) 
      ~ paste0("Tag #", data$TagNo, " (Status ", data$Status,"): Crown Class"),
   is.na(values_data) | data$CrwnCl %in% c("x","X")
      ~ "Crown Class",
   (data$Status == "L" & data$CrwnCl %in% live_classes) | (data$Status == "D" & data$CrwnCl %in% dead_classes)
      ~ paste0("Tag #", data$TagNo, ": Crown Class"),
   TRUE ~ "Crown Class"
)
values_data <- data$CrwnCl
values_valid <- c("D", "C", "I", "SC", "RS", "LBS", "CS", "BAD", "BBD", "DD", "CUS", "O")
values_check <- (values_data %in% values_valid) &
  (data$Status == "L" & values_data %in% live_classes | 
     data$Status == "D" & values_data %in% dead_classes)

#Identify errors
errors_Trees_CrwnCl <- qaqc(data, query, query_message, values_check)
```

Char height values are valid. Rosner's test for outliers, flagging negative values, flagging values that are larger than the height of the tree.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Trees_data %>%
  filter(!is.na(CharHt))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$CharHt)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_CharHt <- min(outliers) - 1

# Set parameters
data <- Trees_data
query <- "Trees Values CharHt"
query_message <- case_when(
    data$CharHt >= data$Ht ~ paste0("Tag #", data$TagNo, ": Char Height > Height"),
    data$CharHt >= max_CharHt ~ paste0("Tag #", data$TagNo, "(Outlier): Char Height"),
    is.na(data$CharHt) ~ "Char Height",
    is.na(data$Ht) ~ "(Height missing): Char Height",
    TRUE ~ paste0("Tag #", data$TagNo, ": Char Height")
    )
    
values_data <- data$CharHt
  # Separate post and non-post
  is_post <- grepl("Post", data$Monitoring.Status)
  is_na <- is.na(values_data)
values_valid <- values_data >= 0 &
                values_data <= max_CharHt &
                values_data <= data$Ht
values_check <- (!is_na & values_valid) |
                (is_na & !is_post)


# Identify errors
errors_Trees_CharHt <- qaqc(data, query, query_message, values_check)
```

Scorch height values are valid. Rosner's test for outliers, flagging negative values, flagging values that are larger than the height of the tree.

```{r}
# Perform Rosner's Test for outliers.
# This test is most appropriate when the sample size is large (n > 20). 

# Filter all data to remove NA values for outlier test
data_temp <- Trees_data %>%
  filter(!is.na(ScorchHt))
# Perform Rosner test on the valid column values
data_rosner <- rosnerTest(data_temp$ScorchHt)
outliers <- data_rosner$all.stats$Value[data_rosner$all.stats$Outlier]
max_ScorchHt <- min(outliers) - 1

# Set parameters
data <- Trees_data
query <- "Trees Values ScorchHt"
query_message <- query_message <- case_when(
    data$ScorchHt >= data$Ht ~ paste0("Tag #", data$TagNo, ": Scorch Height > Height"),
    data$ScorchHt >= max_ScorchHt ~ paste0("Tag #", data$TagNo, "(Outlier): Scorch Height"),
    is.na(data$ScorchHt) ~ "Scorch Height",
    is.na(data$Ht) ~ "(Height missing): Scorch Height",
    TRUE ~ paste0("Tag #", data$TagNo, ": Scorch Height")
    )
values_data <- data$ScorchHt
  # Separate post and non-post
  is_post <- grepl("Post", data$Monitoring.Status)
  is_na <- is.na(values_data)
values_valid <- values_data >= 0 &
                values_data <= max_ScorchHt &
                values_data <= data$Ht
values_check <- (!is_na & values_valid) |
                (is_na & !is_post)


# Identify errors
errors_Trees_ScorchHt <- qaqc(data, query, query_message, values_check)
```

Scorch percent valid values are between 0% and 100%.

```{r}
#Set parameters
data <- Trees_data
query <- "Crown Scorch Percent"
query_message <- "Crown Scorch Percent Value"
values_data <- data$CrScPct
  # Separate post and non-post
  is_post <- grepl("Post", data$Monitoring.Status)
  is_na <- is.na(values_data)
values_valid <- 
  values_data >= 0 &
  values_data <= 100 

values_check <- (!is_na & values_valid) |
                (is_na & !is_post)
  
#Identify errors
errors_Trees_CrScPct <- qaqc(data, query, query_message, values_check)

```

#### Damage Codes:

Codes are valid values

```{r}
#DamCd1
#Set parameters
data <- Trees_data
query <- "Valid Damage Codes"
query_message <- "Damage Code"
values_data <- data$DamCd1
values_valid <- c("ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- 
  is.na(values_data) | 
  (values_data %in% values_valid)

#Identify errors
errors_Trees_DamCd1 <- qaqc(data, query, query_message, values_check)
```

```{r}
#DamCd2
#Set parameters
data <- Trees_data
query <- "Valid Damage Codes"
query_message <- "Damage Code"
values_data <- data$DamCd2
values_valid <- c("ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- 
  is.na(values_data) |
  (values_data %in% values_valid)

#Identify errors
errors_Trees_DamCd2 <- qaqc(data, query, query_message, values_check)
```

```{r}
#DamCd3
#Set parameters
data <- Trees_data
query <- "Valid Damage Codes"
query_message <- "Damage Code"
values_data <- data$DamCd3
values_valid <- c("ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- 
  is.na(values_data) | 
  (values_data %in% values_valid)

#Identify errors
errors_Trees_DamCd3 <- qaqc(data, query, query_message, values_check)
```

```{r}
#DamCd4
#Set parameters
data <- Trees_data
query <- "Valid Damage Codes"
query_message <- "Damage Code"
values_data <- data$DamCd4
values_valid <- c("ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- 
  is.na(values_data) |
  (values_data %in% values_valid)

#Identify errors
errors_Trees_DamCd4 <- qaqc(data, query, query_message, values_check)
```

```{r}
#DamCd5
#Set parameters
data <- Trees_data
query <- "Valid Damage Codes"
query_message <- "Damage Code"
values_data <- data$DamCd5
values_valid <- c("ABGR", "BIRD", "BLIG", "BROK", "BROM", "BURL", "CONK", "CROK", "DTOP", "EPIC", "EPIP", "FIRE", "FORK", "FRST", "GALL", "HOLW", "INSE", "LEAN", "LICH", "LIGT", "MAMM", "MISL", "MOSS", "OZON", "ROOT", "ROTT", "SPAR", "SPRT", "TWIN", "UMAN", "WOND")
values_check <- 
  is.na(values_data) |
  (values_data %in% values_valid)

#Identify errors
errors_Trees_DamCd5 <- qaqc(data, query, query_message, values_check)
```

Combine damage codes for clarity:

```{r}
# Save to Damage Codes error list
errors_Trees_DamCds <- rbind(errors_Trees_DamCd1, errors_Trees_DamCd2, errors_Trees_DamCd3, errors_Trees_DamCd4, errors_Trees_DamCd5)

#Remove all the excess DamCd error data frames
remove(errors_Trees_DamCd1, errors_Trees_DamCd2, errors_Trees_DamCd3, errors_Trees_DamCd4, errors_Trees_DamCd5)
```

#### Checking for duplicate tags:

Tree tags should not have duplicates withing the same plot. Separate by plot and year to differentiate what is really a duplicate.

```{r}
#Set parameters
data <- Trees_data
query <- "Tree Tags"
query_message <- "Duplicate Tree tags"
values_data <- data$TagNo
values_check <-
  data %>%
  group_by(Plot, Date) %>%
  mutate(values_check = duplicated(TagNo) | duplicated(TagNo, fromLast = TRUE)) %>%
  ungroup() %>%
  pull(values_check)
values_check<- !values_check

#Identify errors
errors_Trees_DupTag <- qaqc(data, query, query_message, values_check)
  
```

#### Tree status change

This data check is looking for trees that are "resurrected", meaning they were marked "D" in one read, and then marked "L" in a later read. The error log shows the status of the tree (L/D) and the trees "death date."

Status Columns:

-   Status_Count = How many unique statuses have been recorded for each individual tree? Count of 1 means that the tree has been alive or dead for the entirety of data collection. Count of 2 means that the tree has been both alive and dead.

-   MinDeadDate = What date was the tree first recorded as Dead?

-   MinLiveDate = What date was the tree first recorded as Live?

```{r}
#Add a status count column
Trees_data_CrossYear_Status <- Trees_data %>% 
  group_by(TagNo, MacroPlot.Name) %>% 
  mutate(Status_Count = n_distinct(Status)) %>% 
  ungroup() %>% 
  arrange(TagNo, MacroPlot.Name, Date)
#Status counts will be 1, 2, or 3(only if we have an error in status column)

# add min death date to data_temp
data_temp_Status_AllDead <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count == 1 & Status == "D")
data_temp_Status_SomeDead <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count > 1)
data_temp_Status_Dead <- rbind(data_temp_Status_AllDead, data_temp_Status_SomeDead)

data_temp_MinDeadDate <- data_temp_Status_Dead %>% 
  filter(Status == "D") %>% 
  group_by(TagNo, MacroPlot.Name) %>% 
  mutate(MinDeadDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, MinDeadDate) %>% 
  unique()

data_temp_merge_Status_Dead <- merge(data_temp_Status_Dead, data_temp_MinDeadDate, all = T)

# add min live date to data_temp
data_temp_Status_AllLive <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count == 1 & Status == "L")
data_temp_Status_SomeLive <- Trees_data_CrossYear_Status %>% 
  filter(Status_Count > 1)
data_temp_Status_Live <- rbind(data_temp_Status_AllLive, data_temp_Status_SomeLive)

data_temp_MinLiveDate <- data_temp_Status_Live %>% 
  filter(Status == "L") %>% 
  group_by(TagNo, MacroPlot.Name) %>% 
  mutate(MinLiveDate = min(Date)) %>% 
  select(TagNo, MacroPlot.Name, MinLiveDate) %>% 
  unique()

data_temp_merge_Status_Live <- merge(data_temp_Status_Live, data_temp_MinLiveDate, all = T)

# merge
data_temp_merge_Status <- merge(data_temp_merge_Status_Dead, data_temp_merge_Status_Live, all = T)

remove(Trees_data_CrossYear_Status, data_temp_Status_AllDead, data_temp_Status_SomeDead, data_temp_Status_Dead, data_temp_MinDeadDate, data_temp_merge_Status_Dead, data_temp_Status_AllLive, data_temp_Status_SomeLive, data_temp_Status_Live, data_temp_MinLiveDate, data_temp_merge_Status_Live)
```

Status Dead to Live:

Trees change status from D to L over time.

Filter for Status = L. Check that the sample event date is not later than the death date.

NOTE: Error log will consolidate errors by Tag Number and Plot. This is needed because it flags every instance of the specific tree, not just the read where it became alive again. Functionally, it still achieves the desired objective and the consolidation code cleans up the error log data frame, but it does not highlight the specific year in which the tree "resurrected". The "MinDeadDate" is the starting point. Usually the very next read is where the tree was marked as "Live" once more, but this may not be the case across all plots and reads. For future users: ignore the number in parentheses.

```{r}
# Set parameters 
#data <- merge(Trees_data, data_temp_merge_Status, all = T) %>% 
  #filter(Status == "L")
data <- (data_temp_merge_Status) %>% filter(Status == "L") 
query <- "Trees Status" 
query_message <- paste0("Tag ", data$TagNo, ": Status = ", data$Status, ". ", "Death date")
values_data <- (data$MinDeadDate)
values_valid <- (data$Date)
values_check <- (values_data >= values_valid) %>% replace_na(TRUE)

# Identify errors
errors_Trees_CrossYear_Status <- qaqc(data, query, query_message, values_check)
```

## Trees Error Log

```{r}
# Save to master error list
errors_Trees <- rbind(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia, errors_Trees_FieldTeam, errors_Trees_QTRNum, errors_Trees_PosTag, errors_Trees_Status, errors_Trees_DBH, errors_Trees_Ht, errors_Trees_LiCrBHt, errors_Trees_CrwnCl, errors_Trees_CharHt, errors_Trees_ScorchHt, errors_Trees_CrScPct, errors_Trees_DamCds, errors_Trees_DupTag, errors_Trees_CrossYear_Status)


```

```{r}
#remove all the error data frames
remove(errors_Trees_MacroPlotSize, errors_Trees_SnagPlotSize, errors_Trees_BrkPntDia, errors_Trees_FieldTeam, errors_Trees_QTRNum, errors_Trees_PosTag, errors_Trees_Status, errors_Trees_DBH, errors_Trees_Ht, errors_Trees_LiCrBHt, errors_Trees_CrwnCl, errors_Trees_CharHt, errors_Trees_ScorchHt, errors_Trees_CrScPct, errors_Trees_DamCds, errors_Trees_DupTag, errors_Trees_CrossYear_Status)
```

## Final Exports

```{r}
# Add columns to our data sheets

errors_FuelsFWD_export <- errors_FuelsFWD%>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")

errors_Fuels1000_export <- errors_Fuels1000 %>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")

errors_Fuels_DuffLitt_export <- errors_Fuels_DuffLitt %>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")

errors_PostBurn_export <- errors_PostBurn %>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")

errors_Seedlings_export <- errors_Seedlings %>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")

errors_Trees_export <- errors_Trees %>%
  mutate(
    True_Error = "",
    Validated_by = "",
    Fixed = "",
    Fixed_by = "")


# Save as an excel

write_xlsx(
  list(
    "Fuels FWD" = errors_FuelsFWD_export,
    "Fuels 1000s" = errors_Fuels1000_export,
    "Litter Duff" = errors_Fuels_DuffLitt_export,
    "Post Burn" = errors_PostBurn_export,
    "Seedlings" = errors_Seedlings_export,
    "Trees" = errors_Trees_export
  ),
  path = "/Users/sophiedua/Desktop/SEKI Winter Fellowship/R/2026 SEKI Winter Fellows/output/QAQC_Errors_All_Protocols.xlsx"
)


```
